Supplementary Tables

### 2019-10-11 NAR_form_article.Rmd
typo check

Apart from tools predicting miRNA:target pairs, there are databases that provide  experimentally validated miRNA:target interaction datasets. The reliability in these datasets range from weak (e.g, western blot) to strong (e.g, CLiP, CLASH). High-throughput methods ensure handling of AGO (Argounate Protein) interacted miRNAs (CLiP) and their targets as well (CLEAR-CLiP, CLASH). We constructed a large ceRNA network utilizing high-throughput datasets containing chimeric reading of miRNA:target pairs [@helwak_mapping_2013; @moore_mirnatarget_2015] which expose exact complementary seed regions.

@@

We have analysed the realistic network in context of perturbation efficiencies of nodes. This process was run with 3 fold upregulation and 10 iteration in conditions where elements that have less than five percent change are not taken into account (Codes are found at Section 6.5 in Supplementary Materials and Methods). With this approach, SERPINE2 has been found as one of the most effective gene nodes in network. We have observed that gaining steady-state in realistic network takes a lot of time because of complex interactions, as expected, and SERPINE2 exhibit highest efficiency at iteration 31 with given conditions (see Section 6.6 in Supplementary Materials and Methos and Figure S9 in Supplementary Figures). We ran the model with two fold of iteration which is observed as most effective. We have observed that the results of simulation with SERPINE2 gene, a serine protease inhibitor, one of the overexpressed genes of breast cancer [@yang2018expression; @candia2006protease] is not convenient with tumor tissue expression values, when we compared the tumor gene expression dataset retrieved from TCGA and simulation results obtained via our approach (see Section 6.6 in Supplementary Materials and Methods). Simulation results were not expected to be consistent with tumor tissue expression values because large number of abnormal regulations were observed in these tissues. So, it might not said that a single gene is responsible for entire regulations in cancer cells. Besides, the all other factors such as up/down regulation of miRNAs or other genes were ignored. Therefore, it would be more useful to test our approach with an ideal dataset that includes entire informations of gene expression values before and after the regulation of a known gene at existing tissue, and miRNA expression values of this tissue at initial conditions.

>>>>

Apart from tools predicting miRNA:target pairs, there are databases that provide  experimentally validated miRNA:target interaction datasets. The reliability in these datasets range from weak (e.g, western blot) to strong (.g, CLiP, CLASH). High-throughput methods ensure handling of AGO (Argonaute Protein) interacted miRNAs (CLiP) and their targets as well (CLEAR-CLiP, CLASH). We constructed a large ceRNA network utilizing high-throughput datasets containing chimeric reading of miRNA:target pairs [@helwak_mapping_2013; @moore_mirnatarget_2015] which expose exact complementary seed regions.

@@

We have analysed the realistic network in context of perturbation efficiencies of nodes. This process was run with 3 fold upregulation and 10 iteration in conditions where elements that have less than five percent change are not taken into account (Codes are found at Section 6.5 in Supplementary Materials and Methods). With this approach, SERPINE2 has been found as one of the most effective gene nodes in network. We have observed that gaining steady-state in realistic network takes a lot of time because of complex interactions, as expected, and SERPINE2 exhibit highest efficiency at iteration 31 with given conditions (see Section 6.6 in Supplementary Materials and Methods and Figure S9 in Supplementary Figures). We ran the model with two fold of iteration which is observed as most effective. We have observed that the results of simulation with SERPINE2 gene, a serine protease inhibitor, one of the overexpressed genes of breast cancer [@yang2018expression; @candia2006protease] is not convenient with tumor tissue expression values, when we compared the tumor gene expression dataset retrieved from TCGA and simulation results obtained via our approach (see Section 6.6 in Supplementary Materials and Methods). Simulation results were not expected to be consistent with tumor tissue expression values because large number of abnormal regulations were observed in these tissues. So, it might not said that a single gene is responsible for entire regulations in cancer cells. Besides, the all other factors such as up/down regulation of miRNAs or other genes were ignored. Therefore, it would be more useful to test our approach with an ideal dataset that includes entire informations of gene expression values before and after the regulation of a known gene at existing tissue, and miRNA expression values of this tissue at initial conditions.

### 2019-10-11 Supplementary_additional_file.Rmd
scanning
Figure defining
typo check

MiRNA releases are obtained from [miRBase](http://www.mirbase.org/ftp.shtml). In this step, release 21 (in Human genome 38) is utilised.

@@

CLASH dataset is published in miRBase erlease 15 and Human Genome 19 version.

@@

@@

CLASH dataset was retrieved from [Nature web page](nature.com/articles/ncomms9864)

@@

@@

@@

>>>>

MiRNA releases are obtained from [miRBase](http://www.mirbase.org/ftp.shtml). In this step, release 21 (in Human genome 38) was downloaded.

@@

CLASH dataset is published in miRBase release 15 and Human Genome 19 version.

@@

<!-- #TODO maybe purrr can be used, instead of repeating as.character -->

@@

CLASH dataset was retrieved from [Nature article](nature.com/articles/ncomms9864)

@@

<!-- #TODO the sections above needs some text explaining what is done -->

@@

Numeric values of interaction factors are shown in Table S2-3 in Supplementary Tables file.

@@

The context of dataset is shown in Table S5 in Supplementary Tables.

### 2019-10-11 Supplementary_figures.Rmd
typo check

## *minsamp* dataset analysis at lack of interaction factors.

@@

priming_graph(minsamp, competing_count = Competing_expression, miRNA_count = miRNA_expression)%>%
  update_how("Gene2", 2)%>%
  simulate(cycle=1)%>%
  vis_graph(Competing_color = "navajowhite3", mirna_color = "ivory4", Upregulation = "red", Downregulation = "blue", title = "A")

>>>>

## *minsamp* dataset analysis in absence of interaction factors.

@@

priming_graph(minsamp, 
              competing_count = Competing_expression, 
              miRNA_count = miRNA_expression) %>%
  update_how("Gene2", 2) %>%
  simulate(cycle=1) %>%
  vis_graph(Competing_color = "navajowhite3", 
            mirna_color = "ivory4", 
            Upregulation = "red", 
            Downregulation = "blue", 
            title = "A")

### 2019-10-11 Supplementary_mat_met.Rmd
typo check

The node elements in the dataset are handled as two type; repressive (miRNAs) and competing (targets). It is assumed in approach that while targets are degrading or inhibiting by miRNAs continuously, miRNAs reversibly used. If the trigger of the network is a miRNA, it maintains the current value of amount that provides by user. On the contrary, it tries to help this process to provide steady-state through the regulations on its amount, if a competing element is used as a trigger. The functions that are used in the approach are developed with R programming so as can be used with other packages. These are can be found in the Github repository [ceRNAnetsim Github page](https://github.com/selcenari/ceRNAnetsim) and improved with contributions from others.

@@

When the graphs which were resulted from analyses were examined, it was observed that behaviors were same. But, when the results were analysed in terms of expression values, the regulation differences can be observed.

@@

We downloaded the mirbase release 21 dataset from [mirbase](http://www.mirbase.org/ftp.shtml) and processed the patient mirna expression datasets as following:

@@

perturbationofnetwork%>%
  filter(type== "Competing")%>%
  arrange(desc(perturbation_efficiency))
```
This dataset, perturbationofnetwork, includes totally 423 effective nodes. We selected SERPINE2 gene to perturb the network because it is the most efficient node gene in network.

@@

Actually, we have developed to provide a new approach mirna mediated regulation networks. This approach may not explain the whole regulation behaviors between miRNAs and targets but can be first step to more detailed and coherent miRNA:target regulation approach.

>>>>

The node elements in the dataset are handled as two type; repressive (miRNAs) and competing (targets). It is assumed in approach that while targets are degrading or inhibiting by miRNAs continuously, miRNAs reversibly used. If the trigger of the network is a miRNA, it maintains the current value of amount that provides by user. On the contrary, it tries to help this process to provide steady-state through the regulations on its amount, if a competing element is used as a trigger. The functions that are used in the this study are developed with R and their source code is available in [Github repository](https://github.com/selcenari/ceRNAnetsim).

@@

When the graphs were examined, it was observed that behaviors were same. But, when the results were analysed in terms of expression values, the regulation differences can be observed.

@@

We downloaded the mirbase release 21 dataset from [mirbase](http://www.mirbase.org/ftp.shtml) and processed the patient miRNA expression datasets as following:

@@

perturbationofnetwork%>%
  filter(type== "Competing")%>%
  arrange(desc(perturbation_efficiency))
```
This dataset, `perturbationofnetwork`, includes totally 423 effective nodes. We selected SERPINE2 gene to perturb the network because it is the most efficient node gene in network.

@@

Actually, we have developed to provide a new approach miRNA mediated regulation networks. This approach may not explain the whole regulation behaviors between miRNAs and targets but can be first step to more detailed and coherent miRNA:target regulation approach.

### 2019-10-11 Supplementary_tables.Rmd
scanning
typo check

Table: *minsamp* sample dataset that includes interaction factors.

Note: Energy values in miRNA:target pairs are represented by high-throughput studies [@helwak_mapping_2013; @moore_mirnatarget_2015] which are utilized in this study. On the other hand, we have specified the other interaction factors, seed type and location of binding region on the target, as numeric values based on the previous studies.[@grimson_microrna_2007] have compared the seed types' effect on target repression with few miRNA had canonical seed pairing in their study. Additionally, [@bartel_micrornas:_2009] and [@betel2010comprehensive] have studied on functional and non-functional seed interactions. Based on results of these studies we have arranged seed types of miRNA:target interactions as numeric values. We also have redefined location of binding region on the target as numeric values, based on studies of [@hausser_analysis_2013] and [@helwak_mapping_2013]. With this process, we have handled this entegrated dataset in context of competitor behaviours and functionality of interactions.

@@

 \begin{table}[ht]
\centering
\caption{ miRNA:target pairs supported by High-throughput Experiments}
\begin{tabular}{lll}
  \hline
 Variable & Structure & Means \\
  \hline
   cluster    & character & Barcode from experimentally method \\
   chromosome    & character & Chromosome of Terget gene from raw data \\
   start\_position    & numeric & Gene start position from raw data \\
   end\_position     & numeric & Gene end position from raw data \\
   strand     & chracter & Gene strand \\
   hgnc\_symbol     & character & Gene name (Symbol) \\
   Ensembl\_Gene\_Id     & character & Ensembl Gene Id of gene \\
   Ensembl\_Transcript\_Id   & character & Ensembl transcript id of mRNA of Target gene \\
   target\_seq   & character & mRNA sequences targeted by miRNA \\
   miRNA   & character & miRNA id (from miRBase vesion 21 ) \\
   miR\_seq   & character & miRNA sequence \\
   seed\_type   & character & seed type of miRNA:target interaction \\
   Energy   & numeric & Energy of miRNA:target binding \\
   HG38build\_loc   & character & Recent chromosomal location of Gene \\
   Genom\_build   & character& Genom build of given chromosome, start and end positions \\
   region   & character & interaction location on target \\ 
   region\_effect   & numeric & Coefficient of location on target \\ 
   seed\_type\_effect   & numeric & Coefficient for seed sequence of miRNA:target interaction \\
   \hline
\end{tabular}
\end{table}

As a result of simulation a dataset, a graph object is obtained that includes various variables in edge and node data. A graph object includes variables at followings.

>>>>

Table: *minsamp* sample dataset that includes interaction factors. The parameters which affect miRNA:target interactions (i.e. seed type, region, energy) are provided in sample dataset.

Note: Energy values in miRNA:target pairs are represented by high-throughput studies [@helwak_mapping_2013; @moore_mirnatarget_2015] which are utilized in this study. On the other hand, we have specified the other interaction factors, seed type and location of binding region on the target, as numeric values based on the previous studies.[@grimson_microrna_2007] have compared the seed types' effect on target repression with few miRNA had canonical seed pairing in their study. Additionally, [@bartel_micrornas:_2009] and [@betel2010comprehensive] have studied on functional and non-functional seed interactions. Based on results of these studies we have arranged seed types of miRNA:target interactions as numeric values. We also have redefined location of binding region on the target as numeric values, based on studies of [@hausser_analysis_2013] and [@helwak_mapping_2013]. With this process, we have handled this integrated dataset in context of competitor behaviors and functionality of interactions.

@@

 \begin{table}[ht]
\centering
\caption{ miRNA:target pairs supported by High-throughput Experiments}
\begin{tabular}{lll}
  \hline
 Variable & Structure & Means \\
  \hline
   cluster    & character & Barcode from experimentally method \\
   chromosome    & character & Chromosome of Target gene from raw data \\
   start\_position    & numeric & Gene start position from raw data \\
   end\_position     & numeric & Gene end position from raw data \\
   strand     & character & Gene strand \\
   hgnc\_symbol     & character & Gene name (Symbol) \\
   Ensembl\_Gene\_Id     & character & Ensembl Gene Id of gene \\
   Ensembl\_Transcript\_Id   & character & Ensembl transcript id of mRNA of Target gene \\
   target\_seq   & character & mRNA sequences targeted by miRNA \\
   miRNA   & character & miRNA id (from miRBase version 21 ) \\
   miR\_seq   & character & miRNA sequence \\
   seed\_type   & character & seed type of miRNA:target interaction \\
   Energy   & numeric & Energy of miRNA:target binding \\
   HG38build\_loc   & character & Recent chromosomal location of Gene \\
   Genome\_build   & character& Genome build of given chromosome, start and end positions \\
   region   & character & interaction location on target \\ 
   region\_effect   & numeric & Coefficient of location on target \\ 
   seed\_type\_effect   & numeric & Coefficient for seed sequence of miRNA:target interaction \\
   \hline
\end{tabular}
\end{table}

As a result of simulation a dataset, a graph object is obtained that includes various variables in edge and node data. A graph object includes variables at following.
